---
title: "crosstabr"
author: "Thomas Klebel"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{crosstabr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Basic Usage


# Unused factor levels
`crosstabr` by default preserves unused factor levels. This might be useful
when conducting exploratory tasks -- it would probably bad to miss levels
which were simply not observed.

```{r keep unobserved levels, results='asis'}
library(crosstabr)
test_dat <- titanic # need to export data
index <- test_dat$Class == "2nd"
test_dat[index, ] <- NA

test_dat %>%
  crosstab(Survived ~ Class)
```

If you want to remove unused levels, simply add `droplevels = FALSE` when setting
up the layout:

```{r drop unobserved levels}
test_dat %>%
  crosstab(Survived ~ Class) %>%
  layout_column(droplevels = T)
```

# Programming with crosstabr

You can program with `crosstabr` in several ways.

## Wrapping the pipeline
In case you want to reuse a certain configuration of using `crosstabr`, you can
simplay wrap it in a function:

```{r}
fun <- function(data, formula) {
  data %>%
    crosstab(formula) %>% 
    layout_column() %>% 
    add_stats(chisq.test)
}

fun(titanic, Survived ~ Class)
```


## Creating formula from (string) variables
In case you want to refer to the names of variables via a string variable, you
can use `reformulate` to construct the formula for you:

```{r, message=FALSE}
fun2 <- function(data, dep, indep) {
  data %>%
    crosstab(reformulate(indep, response = dep))
}

dep <- "Survived"
indep <- "Class"

fun2(titanic, dep, indep)
```

